---
title: "p8105_hw3_sc5078"
author: "Yvonne Chen"
date: "2023-10-14"
output: github_document
---

# Problem 0
```{r setup}
library(tidyverse)
library(ggridges)
library(patchwork)


library(p8105.datasets)

```

# Problem 1
load the dataset, instacart
```{r import "instacart"}
data("instacart")
```
This dataset comprises `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row corresponds to a single product within an Instacart order. The variables encompass user, order, and product identifiers, as well as the order placement sequence. Order-level attributes include details about the order's day, time, and days since the previous order. Additionally, item-specific characteristics cover the product name (e.g., Yogurt, Avocado), department (e.g., dairy and eggs, produce), aisle (e.g., yogurt, fresh fruits), and whether the item has been previously ordered by the user. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` distinct products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` unique orders placed by `r instacart |> select(user_id) |> distinct() |> count()` different users.


This is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables being the most items ordered by far.
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```


The plot shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```


This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```


This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


# Problem 2
load the dataset, BRFSS
```{r import "BRFSS"}
data("brfss_smart2010")
```

data cleaning
```{r "BRFSS" data cleaning}
brfss_smart2010 = brfss_smart2010 |>
  janitor::clean_names()|>
  filter(topic %in% "Overall Health") |>
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
  mutate(response = ordered(response, levels=c("Poor", "Fair", "Good", "Very good", "Excellent"))) |>
  arrange(response)
```


List of states were observed at 7 or more locations in year 2002 and 2010.
```{r}
brfss_smart2010 |>
  filter(year %in% c(2002, 2010)) |>
  group_by(year, locationabbr) |>
  filter(n_distinct(locationdesc) >= 7) |>
  summarize()
```


Construct a dataset, excellent, that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. This is a “spaghetti” plot of this average value over time within a state.
```{r}
excellent <- brfss_smart2010 |>
  filter(response %in% "Excellent") |>
  group_by(year, locationabbr) |>
  mutate(
    average_data = mean(data_value)
  ) |>
  summarize(year, average_data)

ggplot(excellent, aes(x = year, y = average_data, color = locationabbr)) +
  geom_line()  +
  labs(title = "Average Value Over Time within a State",
       x = "Year",
       y = "Average Value",
       color = "State") + 
  theme_minimal()
```


There is a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_smart2010 |>
  filter(year %in% c(2006, 2010)) |>
  filter(locationabbr %in% "NY") |>
  ggplot(aes(x = response, y = data_value, fill = response)) +
  geom_boxplot() +
  facet_wrap(~year, nrow = 1) +
  labs(title = "Distribution of Data Value for Responses in NY State",
       x = "Response",
       y = "Data Value") +
  theme_minimal()
```


# Problem 3
Load and tidy dataset, nhanes_accel
```{r load and tidy dataset, nhanes_accel}
nhanes_accel = read.csv("./data/nhanes_accel.csv") |>
  janitor::clean_names()|>
  pivot_longer(
    min1:min1440,
    names_to = "minute", 
    names_prefix = "min",
    values_to = "min") |>
  mutate(
    minute = as.numeric(minute),
    min = as.numeric(min)
  )
```

Load and tidy dataset, nhanes_covar
```{r load and tidy dataset, nhanes_covar}
nhanes_covar = read.csv("./data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names()|>
  na.omit() |>
  filter(age >= 21) |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"), 
    sex = as.factor(sex),
    education =
      case_match(
        education,
        1 ~ "Less than high school",
        2 ~ "High school equivalent",
        3 ~ "More than high school"
      ),
    education = as.factor(education)
  )
```

Merge nhanes_accel and nhanes_covar into nhanes
```{r merge to nhanes}
nhanes = 
  left_join(nhanes_covar, nhanes_accel, by = "seqn")
```

This is a reader-friendly table for the number of men and women in each education category.
```{r}
nhanes |>
  janitor::tabyl(education, sex)
```
There are more male (50400) have high school equivalent education level than female (33120). There are more female have less than high school education or more than high school education than male.

Create a visualization of the age distributions for men and women in each education category.
```{r}
nhanes |>
  group_by(education, sex) |>
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions by Education Category",
       x = "Education Category",
       y = "Age",
       fill = "Sex")

```
The age distribution between female and male is similar in the "less than high school" and "More than high school" education group. However, for the "High school equivalent" education level, females would have a higher maximum age, minimum age, and average age than males. For females, "High school equivalent" education level has the highest mean age, `r filter(nhanes, sex == "female", education == "High school equivalent") |> pull(age) |> mean() |> round(2)`, while "More than high school" education level has the lowest mean age, `r filter(nhanes, sex == "female", education == "More than high school") |> pull(age) |> mean() |> round(2)`. For males, "Less than high school" have the highest mean age, `r filter(nhanes, sex == "male", education == "Less than high school") |> pull(age) |> mean() |> round(2)`, while "More than high school" education level has the lowest mean age, `r filter(nhanes, sex == "male", education == "More than high school") |> pull(age) |> mean() |> round(2)`.


Aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis).
```{r}
nhanes |>
  group_by(seqn, age, sex, education) |>
  summarize(total_activity = sum(min)) |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(group = sex), linetype = 2) +
  facet_wrap(~ education, nrow = 2) +
  labs(title = "Total Activity vs. Age by Education Level",
       x = "Age",
       y = "Total Activity",
       color = "Sex") +
  theme_minimal()
```




