---
title: "p8105_hw3_sc5078"
author: "Yvonne Chen"
date: "2023-10-14"
output: github_document
---

# Problem 0
```{r setup}
library(tidyverse)
library(ggridges)
library(patchwork)


library(p8105.datasets)

```

# Problem 1
load the dataset, instacart
```{r import "instacart"}
data("instacart")
```
This dataset comprises `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row corresponds to a single product within an Instacart order. The variables encompass user, order, and product identifiers, as well as the order placement sequence. Order-level attributes include details about the order's day, time, and days since the previous order. Additionally, item-specific characteristics cover the product name (e.g., Yogurt, Avocado), department (e.g., dairy and eggs, produce), aisle (e.g., yogurt, fresh fruits), and whether the item has been previously ordered by the user. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` distinct products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` unique orders placed by `r instacart |> select(user_id) |> distinct() |> count()` different users.


This is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables being the most items ordered by far.
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```


The plot shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```


This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```


This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


# Problem 2
load the dataset, BRFSS
```{r import BRFSS}
data("brfss_smart2010")
```

data cleaning
```{r BRFSS data cleaning}
brfss_smart2010 = brfss_smart2010 |>
  janitor::clean_names()|>
  filter(topic %in% "Overall Health") |>
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
  mutate(response = ordered(response, levels=c("Poor", "Fair", "Good", "Very good", "Excellent"))) |>
  arrange(response)
```


List of states were observed at 7 or more locations in year 2002 and 2010.
```{r}
brfss_smart2010 |>
  filter(year %in% c(2002, 2010)) |>
  group_by(year, locationabbr) |>
  filter(n_distinct(locationdesc) >= 7) |>
  summarize()
```


Construct a dataset, excellent, that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. This is a “spaghetti” plot of this average value over time within a state.
```{r}
excellent <- brfss_smart2010 |>
  filter(response %in% "Excellent") |>
  group_by(year, locationabbr) |>
  mutate(
    average_data = mean(data_value)
  ) |>
  summarize(year, average_data)

ggplot(excellent, aes(x = year, y = average_data, color = locationabbr)) +
  geom_line()  +
  labs(title = "Average Value Over Time within a State",
       x = "Year",
       y = "Average Value",
       color = "State") + 
  theme_minimal()
```


There is a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_smart2010 |>
  filter(year %in% c(2006, 2010)) |>
  filter(locationabbr %in% "NY") |>
  ggplot(aes(x = response, y = data_value, fill = response)) +
  geom_boxplot() +
  facet_wrap(~year, nrow = 1) +
  labs(title = "Distribution of Data Value for Responses in NY State",
       x = "Response",
       y = "Data Value") +
  theme_minimal()
```


# Problem 3
Load and tidy dataset, nhanes_accel
```{r load and tidy dataset, nhanes_accel}
nhanes_accel = read.csv("./data/nhanes_accel.csv") |>
  janitor::clean_names()|>
  pivot_longer(
    min1:min1440,
    names_to = "minute", 
    names_prefix = "min",
    values_to = "mims") |>
  mutate(
    minute = as.numeric(minute),
    mims = as.numeric(mims)
  )
```

Load and tidy dataset, nhanes_covar
```{r load and tidy dataset, nhanes_covar}
nhanes_covar = read.csv("./data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names()|>
  na.omit() |>
  filter(age >= 21) |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"), 
    sex = as.factor(sex),
    education =
      case_match(
        education,
        1 ~ "Less than high school",
        2 ~ "High school equivalent",
        3 ~ "More than high school"
      ),
    education = as.factor(education)
  )
```

Merge nhanes_accel and nhanes_covar into nhanes
```{r merge to nhanes}
nhanes = 
  left_join(nhanes_covar, nhanes_accel, by = "seqn")
```

This is a reader-friendly table for the number of men and women in each education category.
```{r}
nhanes |>
  group_by(sex, education)|>
  summarize(n_obs = n_distinct(seqn))|>
  pivot_wider(
    names_from = education,
    values_from = n_obs
  )|>
  knitr::kable()
```
There are more male having high school equivalent education level than female. There are more female have less than high school education or more than high school education than male. For both male and female, most of people having more than high school education.

Create a visualization of the age distributions for men and women in each education category.
```{r}
nhanes |>
  group_by(education, sex) |>
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions by Education Category",
       x = "Education Category",
       y = "Age",
       fill = "Sex")

```

The age distribution between female and male is similar in the "less than high school" and "More than high school" education group. However, for the "High school equivalent" education level, females would have a higher maximum age, minimum age, and average age than males. For females, "High school equivalent" education level has the highest mean age, `r filter(nhanes, sex == "female", education == "High school equivalent") |> pull(age) |> mean() |> round(2)`, while "More than high school" education level has the lowest mean age, `r filter(nhanes, sex == "female", education == "More than high school") |> pull(age) |> mean() |> round(2)`. For males, "Less than high school" have the highest mean age, `r filter(nhanes, sex == "male", education == "Less than high school") |> pull(age) |> mean() |> round(2)`, while "More than high school" education level has the lowest mean age, `r filter(nhanes, sex == "male", education == "More than high school") |> pull(age) |> mean() |> round(2)`.


Aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis).
```{r}
nhanes |>
  group_by(seqn, age, sex, education) |>
  summarize(total_activity = sum(mims)) |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE, aes(group = sex), linetype = 2) +
  facet_wrap(~ education, nrow = 2) +
  labs(title = "Total Activity vs. Age by Education Level",
       x = "Age",
       y = "Total Activity",
       color = "Sex") +
  theme_minimal()
```

For people with high school equivalent education level, female and male have similar shape of distribution of total activity vs. age, but female tends to have higher total activity at all age levels comapred to male. Both female and male are likely to have the highest total activity at around age 40; both female and male are likely to have the lowest total activity at around age 20, 60, and 80. Total activity increase from age 20 to 40, decrease from age 40 to 60, increase from age 60 to 70, and decrease from age 70 to 80.

For people with less than high school education, female and male have similar distribution of total activity vs. age. From age 20 to 40, female has higher total activity than male; from age 40 to 80, males has higher total activity than females. The highest total activity for both female and male lies around age 20, and have a generally decreasing pattern through time. The lowest total activity for both female and male lies around age 80.

For people with more than high school education, female has a steady distribution of total activity from age 20 to 60, and decreases from 60 to 80; males has a steady distribution of total activity from age 20 to 50, and decrease from 50 to 80. Generally, male has a lower total activity compared to female at all ages.

Generally speaking, people with less than high school education has the largest total activity, but people with more than high school education tends to have the most stable total activity through time, especially for females. However, it's still hard to conclude the relationship between education level and total activity for different sex groups.


Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.
```{r}
nhanes |>
  group_by(sex, education, minute) |>
  summarise(mims_mean = mean(mims)) |>
  ggplot(aes(x = minute, y = mims_mean, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE, aes(group = sex), linetype = 2) +
  facet_wrap(~ education, nrow = 2) +
  labs(title = "24-hour activity time courses by Education Level",
       x = "Minute",
       y = "Mims",
       color = "Sex") +
  theme_minimal()

```

For people with high school equivalent education level, female and male have generally same distribution of activity vs. time while female tends to have slightly higher activity than male. Both female and male decreases activity from 0 to 250 minutes, increase from 250 to 500 minutes, become most active during 500 to 1000 minutes, and decrease until 1440 minutes.

For people with less than high school education, female and male have generally same distribution of activity vs. time while female tends to have slightly higher activity than male. Both female and male decreases activity from 0 to 250 minutes, increase from 250 to 700 minutes (arriving the most active time), and decrease from 700 to 1440 minutes.

For people with more than high school education, female and male have generally same distribution of activity vs. time while female tends to have slightly higher activity than male. Both female and male decreases activity from 0 to 250 minutes, increase from 250 to 500 minutes, become most active during 500 to 1200 minutes, and decrease from 1200 to 1440 minutes.

Generally speaking, people with more than high school education have the largest time period of being most active, the second most time of being active is people with high school or equivalent education, and people with less than high school education have the least time of being most active. It's fair to concludes from the plots that with higher education level, people tends to keep active for a longer time period.
